SKETCH RECOGNITION AND UNDERSTANDING

### 2009-10-13 draft/System.tex
FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL

Two DPSO algorithms generate two different segmentations for each stroke. Each DPSO algorithms tries to find the best particle in the solution space. Both algorithms has the same problem formulation but different fitness functions. After generating two segmentations the system chooses the best segmentation from the outputs of the DPSO. The details of the problem formulation and fitness functions are given in the following sections.

@@

The system represents each particle $P_i$ by $P_i = \left\{ {p_{ij} \left| {j = 1,2 \cdots N} \right.} \right\}$ where $p_{ij}$ is either 0 or 1 where $p_{ij}=1$ means that point $j$ is a dominant point(see Figure \ref{fig:CodingSwarm}.  Thus , the particle represents points that are chosen to be used as dominant points for this segmentation. 

@@

The second algorithm has the same problem formulation but different fitness and error functions. It was previously introduced in  \cite{CruveDivisionSwarm} where genetic programming was used as the optimizing algorithm. The particle is represented using the arrays $P_i = \left\{ {p_{ij} \left| {j = 1,2 \cdots N} \right.} \right\}$ where $p_{ij}$ is either 0 or 1. Lets denote the segments $\widehat{x_ix_j}$ as the consecutive set of points from point $x_i,x_{i+1} \cdots,x_j$. The algorithms attempts to fit each segment $\widehat{x_ix_j}$ to both line and circle. The type of the segment is determined according to the minimum error of each fit. The final error of the solutions is the summation of error of all segments.  %The goal of the algorithm is to fit the segments between the solution vertices' into straight lines or a circular arc. 
\paragraph{Fitting segments into straight line.}
\label{sec:FittingSegmentIntoStraightLine}
According to \cite{CruveDivisionSwarm} each segment $\widehat{x_ix_j}$ given a set of points $x_i,x_{i+1} \cdots,x_j$ can be fitted into the line: $y=kx+c$ where $k$ and $c$ are the slope and the intercept of the line respectively. Equation \ref{eq:Linek} and \ref{eq:LineC} 
\begin{equation}
\label{eq:Linek}
k = \frac{{N\sum\limits_{i = 1}^N {x_i y_i }  - \sum\limits_{i = 1}^N {x_i } \sum\limits_{i = 1}^N {y_i } }}{{N\sum\limits_{i = 1}^N {x_i^2 }  - \left( {\sum\limits_{i = 1}^N {x_i } } \right)^2 }}
\end{equation}
\begin{equation}
\label{eq:LineC}
c = \frac{{\sum\limits_{i = 1}^N {x_i^2 } \sum\limits_{i = 1}^N {y_i }  - \sum\limits_{i = 1}^N {x_i } \sum\limits_{i = 1}^N {x_i y_i } }}{{N\sum\limits_{i = 1}^N {x_i^2 }  - \left( {\sum\limits_{i = 1}^N {x_i } } \right)^2 }}
\end{equation}
, where $N$ is the number of points in the segment and $(x_i, y_i)$ are the coordinate of the point $i$, are used to fit the segment $\widehat{x_ix_j}$ into the straight line.
\begin{equation}
\label{eq:ds}
 d_s  = \frac{{\sum\limits_{i = 1}^N {\left| {(kx_i  + c) - y_i } \right|} }}{{N\sqrt {k^2  + 1} }}
\end{equation}
The distance ($d_s$) (Equation \ref{eq:ds}) is the average distance from segment points $(x_i, y_i)$ to the estimated line. This distance $d_s$ is used as the error of the line approximation.
\paragraph{Fitting segments into circular arc.}
\label{sec:FittingSegmentcirculararc}
Each segment is also fitted into a circular arc $(x - a)^2  + (y - b)^2  = R^2, $ where ($a,b$) are the coordinate of the center of the circle and $R$ is the radius of the circle. They is estimated using the following set of equations:

@@

\subsection{Choosing the Best Fit}
\label{sec:bestFit}
The two PSO algorithms produce two segmentation solutions. After that the system evaluates each solution to finally select the best segmentation. This evaluation is based on the area of both the segmentation and the input stroke. The segmentation error($E_{alg}$ )represents the sum of Square Root Error of the length from the point on the stroke to the estimated curve or line. In \textsl{AlgS2}, this error is the same as the computed error in Equation \ref{eq:ErrorSwarm1} but For the second algorithm \textsl{AlgS2} it is different than the computed error in Equation \ref{eq:errorSwarm2}. The additional computation is necessary to  compare the two segmentation fairly. The system choose the minimum segmentation as the final segmentation of this stroke.

After the best fit segmentation is chosen the parameter of each primitive is computed. For example, if the stroke is represented as an ellipse, the center and axis of the ellipse is computed and confirmed. For lines the slope, equation, start point and end point are computed.  These values and parameter will help determine the spatial relations between different segments in the feature extraction step. 
% The segmentation which crossponds to the minimum $E_{tot}$ is the best segmentation.  
%\begin{equation}
%\label{eq:bestFit1}
%E_{area}=A_{st}-A_{sg}
% \end{equation}
% \begin{equation}
% where $A_{st}$ is the area of the input stroke, $A_{sg}$ is area of computed segmentation, $E_{alg}$ is the Segmentation Error.

@@

After the segmentation, each stroke is represented as a list of segments $L_s$. A segment can be either an ellipse, a line or a circular arc. The list of segments $L_s$ represents the segmentation of the stroke. When a user draw another stroke the generated segmentation of the stroke is appended to the list of segments.  When the user finish drawing the symbol the segment list $L_s$ will contain all segments of the drawn symbol. The relation and properties between different segments are extracted from the segment list $L_s$.  This step is considered as feature extraction which is described in details in the next sections.

@@

\subsection{Structural and geometrical Features(FS1)}
 Features defines the structure of the geometrical symbol. Section \ref{sec:steprec} describe in details how these 13 feature are computed.  
%\begin{description}
% \item [Structural and geometrical Features (FS1)] 
  \begin{itemize}
	 \item \emph{Segments:} Number of segments in the symbol.
	 \item \emph{Strokes:} Number of strokes or partial strokes that created the symbols.  
		\item  \emph{Primitives:} Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
		\item \emph{Curves:} Number of curves or ellipses in the symbol. 
		\item \emph{Lines:} Number of lines in the symbol. 
		\item \emph{Perpendicular} lines Number of perpendicular lines. 
		\item \emph{Parallel} lines Number of parallel lines. 
		\item \emph{Intersections} Number of intersection between lines and curves. 
		\item \emph{T intersections} Number of T intersections. 
		\item \emph{L intersections} Number of L intersections. 
		\item \emph{X intersections} Number of X intersections.
		\item \emph{Min Radius} Minimum Radius of all curves in the symbol.
		\item \emph{Max Radius} Maximum Radius of all curves in the symbol.
		%\item[No. of holes] Number of holes in the symbols. 
\end{itemize}
%\begin{description}
%	 \item[Segments] Number of segments in the symbol.
%	 \item [Strokes] Number of strokes or partial strokes that created the symbols.  
%		\item[Primitives] Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
%		\item [Curves] Number of curves or ellipses in the symbol. 
%		\item [Lines] Number of lines in the symbol. 
%		\item [Perpendicular lines] Number of perpendicular lines. 
%		\item [Parallel lines] Number of parallel lines. 
%		\item [Intersections] Number of intersection between lines and curves. 
%		\item [T intersections] Number of T intersections. 
%		\item [L intersections] Number of L intersections. 
%		\item [X intersections] Number of X intersections.
%		
%		
%		%\item[No. of holes] Number of holes in the symbols. 
%
%\end{description}
\subsection{Rubine Feature Set (FS2)} Features introduced by Rubine\cite{gestureexample12} for single stroke gestures. These 12 features represent the global shape of the symbol drawn. To compute the features we append all segments points into single path $ink_{path}$. The features is computed on the generated path $ink_{path}$.%then compute the features based on this path.  
\begin{enumerate}
	\item Cosine of starting angle.
	\item Sine of starting angle.
	\item Length of diagonal of bounding box. It gives an idea of the size of the bounding box).
	\item Angle of diagonal. It gives an idea of the shape of the bounding box (long, tall, square).
	\item Distance from start to end.  
	\item Total stroke length
	\item Change in Rotation( Arctan ). It gives the directional angle.
	\item Absolute rotation 
	\item Rotation squared 
	\item The maximum speed reached (squared) 
	\item Total time of stroke 
\end{enumerate}
\subsection{Statistical Features (FS3)} 
\begin{enumerate}
	\item \textsl{Zernike moments:} Zernike moments of order $n$\footnote{ We tried different values of n. Results show the analysis of each order.} \cite{HeloiseBeautification}.  For $n=10$ there are 32 moment features. 
\end{enumerate}
\subsection{Global shape properties set (FS4)}
 Features that are composites of different geometrical and statistical symbol. These 13 features represent the global shape of the symbol drawn.

@@

\section{Classification}%%Feature Set
\label{sec:Classification}
The system uses support vector machine (SVM) classifier with Linear kernel (Linear kernel) \cite{libsvm}. SVM is a binary classifier that maximizes the margin between the decision boundary and the support vectors. To enable SVM to learn multi classes we used one versus one strategy for combining set of binary classifier. This strategy trains $ \frac{n(n-1)}{{2}}$ binary classifier each distinguish between only two shapes. The final result is based on a number of votes each shape has from all classifiers. We used the library implementation of LibSVM \cite{libsvm} and its implementation of the one-vs-one strategy. Section \ref{Sec:SVMdetail} describe in details the classifier, how classes (classification) are determined and the reasons behind using this specific classification methods.

@@

\subsection{Support Vector Machines}
\label{Sec:SVMdetail}
The foundation of Support Vector Machines (SVM) were developed by Vapnik \cite{svmintroduce} in 1995. It is gaining increase popularity due to the high performance and efficient implementation. The goal of SVM algorithms is to generate a separating plane between two different classes which are represented using a set of n-dimensional vectors. The generated plane is used as a good classifier to the unseen examples. However, there can be more than one plane that can separates the data therefore SVM attempts to find the plane that maximize the margin (distance between classifier and nearest sample) between the two classes \ref{fig:svm1}. To calculate the margin, two different hyper planes are constructed, one plane on each side of the data, these two planes are then moved to generate the maximum margin possible.  
\begin{figure}
	\centering
		\includegraphics{afterDefense/svm1.jpg.eps}
	\caption{Different Classifier Planes }
	\label{fig:svm1}
\end{figure}

@@

 The orientation of the separation plane is determined using the vector $\omega$ and the offset of the plane from the origin is determined using the scalar $b$.   Let's make an assumption that the data point are linearly separable, i.e. there are infinite possibilities to find a plane that separates and correctly classify the training data. This is idea is illustrated in Figure\ref{fig:svm1} where two different separating planes can correctly classify the data. It is intuitively noted from the figure that the solid line is more likely to generalize more data. This means less classification error will be introduced. The geometrical characterization of this plane means the plane furthest from both classes.     % Which one is preferable ? Intuitively one prefers the solid pane since small perturbations of any point would not introduce misclassifications errors. Without any information, the solid plane is more likely to generalize better on future data. Geometrically we can characterize the solid plane as being furtherer from both classes.  Plane supports a class if all points in that class are on one side of that plane.

@@

\subsubsection{Linearly Inseparable Case}
In all previous sections we assumed that the problem of classifying two set is linearly separable. The assumption is not true in most cases, thus, trying to construct a plane to separate the two dataset will fail. Figure \ref{fig:svm5} illustrates that we cannot construct a plane that can separate the two sets correctly. The constrains for constructing the planes must be relaxed since the QP task is not feasible for the linearly inseparable case.  Consider the linear inseparable problem in Figure\ref{fig:svm5}, in the ideal case there will be no point misclassified and no point falls on the margin. But we must relax the constrain that insure that each point in on the appropriate side of it is supporting plane. Every point that falls on the wrong side of the supporting plane is considered an error. The problem is now to maximize the margin and minimize the error.

@@

\subsubsection {Non linear function Using Kernels}
Figure \ref{fig:svm6} shows a classification problem that cannot be separated using linear space. A higher order plane can work better in similar cases. Converting a linear classification algorithm into non linear classification algorithm can be achieved by adding additional attribute to the data that are nonlinear functions of the original data. The expanded dataset features space can be separated using an existing linear classification algorithm.  The linear classification that is applied on the feature space produces a non linear classification on the original input space. In two dimensional vector space with attributes $r$ and $s$, to construct a quadratic  discriminate  simply map the original two dimensional input space $\left[r,s\right]$ to  five dimensional features space $\left[r,s,rs,r^2,s^2\right]$ and construct a linear discriminate in that space. specifically, define : $\vartheta(x):R^2  \rightarrow R^5$ then 
\begin{equation}
\begin{array}{l}
	x=[r,s] \\
	\omega \cdot x=\omega_1 r+ \omega_2 s \\
	\downarrow \\
	\vartheta(x)=\left[r,s,rs,r^2,s^2\right] \\

@@

 There is mainly two popular methods to change the binary SVM classification problem into Multi class classification problem. The first methods is One Versus All (OVA) Strategy where M binary classifiers is constructed one classifier for each class.  The $ith$ classifier is trained using positive example form $ith$ class and negative example from all other classes. The output function $p_i$ represents the distance and classification of the example from the $ith$ classifier. For a new example $x$, OVA SVM strategy assigns it to the class with the largest value of $p_i$.

@@

SVM have been applied to the a largest MNIST a Handwritten digit recognition dataset of 60000 training set and 10000 test set\cite{IjdarSherifPaper}. Sherif and Ezzat has performed a comparison between SVM and several other techniques on the MNIST dataset. They also compared the same classification methods on Arabic handwritten digits (MADBASE) \cite{ADBase9,IjdarSherifPaper}. Their results shows that SVM classifier outperform all other techniques on both MNIST and MADBase datasets and achieved recognition rate of over 98\%. Similar results was reported by \cite{empiricalcomp11,SVMInvariantComDecoste02} on the MNIST dataset. There are many other successful application SVM. They have proven to be robust to noise and perform well on many tasks \cite{empiricalcomp11,Scholkopf97comparingsupport}.

This superior performance because they may eliminates many of problem experienced with other inference methodologies with other inference methodologies like neural network and decision trees. The first problem that is eliminated was the problems with local minima. SVM can construct highly nonlinear classification and regression function without working about getting stuck at local minima. Another problem SVM eliminated is the numbers of parameter that are picked to generate the model. For example, SVM classifier with RBF kernel will only need 2 parameters to set the model the $C$ cost parameter and $\sigma$ parameter of the RBF kernel. The last reason is the final result are sable, reproducible and largely independent of the specific algorithm used to  optimize the SVM model. If two users apply the same SVM model with the same parameters to the same data, they will et the same solution module numeric issues. Compare this with neural networks where the results are dependent on the particular algorithms and starting point used.

@@

\subsubsection {SVM Implementations}
To solve SVM approach requires solving a QP or LP problem.  Mathematical programming have extensively studied the LP and QP type problems. These techniques can be divided into three categories, methods in which evaluate and discard kernel components during learning. Decomposition methods in which evolving subsets of data are used and new optimization approaches that specifically utilize the structure of the SVM problem.

@@

The third approach is to directly attack the SVM problem from and optimization perspective and create algorithms that explicitly exploits the structure of ten problem. Keerthi et al. \cite{fastSVM6} proposed a very effective algorithm based on the dual geometry of finding the two closest points in the convex hulls.

@@

In this step the user draw one or more stroke to draw a symbol. Figure \ref{fig:clock} illustrate the sequence and direction of user draw. In this section, each stroke user draws will be labeled as it illustrated in Figure \ref{fig:clockstroke1} and Figure \ref{fig:clockstroke2}.  Table \ref{tab:StepsStroke1} shows the detailed output of the program at each step. The input of this step is the a set of strokes, each stroke has a set of  points and the output is a list of segments that will be used in the recognition process. More detailed example and algorithms are demonstrated in the Appendix \ref{ChapterstepExample} \\

@@

	\includegraphics[scale=0.5]{afterDefense/clockstroke1.eps}
  &   &    
\\ \hline 
Segmentation & Segmentation of Ellipse & $P_{dp}$   &  Ellipse fitting  & 
The circle is   a = 103.05715108084719 ,b = 103.06851874747797 , Center( 283.03365107691775,  181.05035287667172 )   with Error =  1.9471676673301757 , Percent =1.0695193086531243
eff  = 0.27463462099275116

@@

Stroke 2 & Extracting point and computing boundary data & -- &  see Section \ref{sec:Preprocessing} &
 	\includegraphics[scale=0.5]{afterDefense/clockstroke2.jpg.eps} \\ \hline
  & Possible Dominate point Extraction for details see Section\ref{sec:Preprocessing} &   	\includegraphics[scale=0.5]{afterDefense/clockstroke2.jpg.eps}  & 1. compute the $P_{pd}$  & 
    \\ \hline

>>>>

Two DPSO algorithms generate two different segmentations for each stroke. Each DPSO algorithms tries to find the best particle in the solution space. Both algorithms have the same problem formulation but different fitness functions. After generating two segmentations the system chooses the best segmentation from the outputs of the DPSO. The details of the problem formulation and fitness functions are given in the following sections.

@@

The system represents each particle $P_i$ by $P_i = \left\{ {p_{ij} \left| {j = 1,2 \cdots N} \right.} \right\}$ where $p_{ij}$ is either 0 or 1 where $p_{ij}=1$ means that point $j$ is a dominant point(see Figure \ref{fig:CodingSwarm}.  Thus, the particle represents points that are chosen to be used as dominant points for this segmentation. 

@@

The second algorithm has the same problem formulation but different fitness and error functions. It was previously introduced in  \cite{CruveDivisionSwarm} where genetic programming was used as the optimizing algorithm. The particle is represented using the arrays $P_i = \left\{ {p_{ij} \left| {j = 1,2 \cdots N} \right.} \right\}$ where $p_{ij}$ is either 0 or 1. Lets denote the segments $\widehat{x_ix_j}$ as the consecutive set of points from point $x_i,x_{i+1} \cdots,x_j$. The algorithms attempts to fit each segment $\widehat{x_ix_j}$ to both line and circle. The type of the segment is determined according to the minimum error of each fit. The final error of the solutions is the summation of error of all segments.  %The goal of the algorithm is to fit the segments between the solution vertices' into straight lines or a circular arc. 
\paragraph{Fitting segments into straight line.}
\label{sec:FittingSegmentIntoStraightLine}
According to \cite{CruveDivisionSwarm} each segment $\widehat{x_ix_j}$ given a set of points $x_i,x_{i+1} \cdots,x_j$ can be fitted into the line: $y=kx+c$ where $k$ and $c$ are the slope and the intercept of the line respectively. Equation \ref{eq:Linek} and \ref{eq:LineC} 
\begin{equation}
\label{eq:Linek}
k = \frac{{N\sum\limits_{i = 1}^N {x_i y_i }  - \sum\limits_{i = 1}^N {x_i } \sum\limits_{i = 1}^N {y_i } }}{{N\sum\limits_{i = 1}^N {x_i^2 }  - \left( {\sum\limits_{i = 1}^N {x_i } } \right)^2 }}
\end{equation}
\begin{equation}
\label{eq:LineC}
c = \frac{{\sum\limits_{i = 1}^N {x_i^2 } \sum\limits_{i = 1}^N {y_i }  - \sum\limits_{i = 1}^N {x_i } \sum\limits_{i = 1}^N {x_i y_i } }}{{N\sum\limits_{i = 1}^N {x_i^2 }  - \left( {\sum\limits_{i = 1}^N {x_i } } \right)^2 }}
\end{equation}
, where $N$ is the number of points in the segment and $(x_i, y_i)$ are the coordinate of the point $i$, are used to fit the segment $\widehat{x_ix_j}$ into the straight line.
\begin{equation}
\label{eq:ds}
 d_s  = \frac{{\sum\limits_{i = 1}^N {\left| {(kx_i  + c) - y_i } \right|} }}{{N\sqrt {k^2  + 1} }}
\end{equation}
The distance ($d_s$) (Equation \ref{eq:ds}) is the average distance from segment points $(x_i, y_i)$ to the estimated line. This distance $d_s$ is used as the error of the line approximation.
\paragraph{Fitting segments into circular arc.}
\label{sec:FittingSegmentcirculararc}
Each segment is also fitted into a circular arc $(x - a)^2  + (y - b)^2  = R^2, $ where ($a,b$) are the coordinate of the center of the circle and $R$ is the radius of the circle. They are estimated using the following set of equations:

@@

\subsection{Choosing the Best Fit}
\label{sec:bestFit}
The two PSO algorithms produce two segmentation solutions. After that the system evaluates each solution to finally select the best segmentation. This evaluation is based on the area of both the segmentation and the input stroke. The segmentation error ($E_{alg}$ )represents the sum of Square Root Error of the length from the point on the stroke to the estimated curve or line. In \textsl{AlgS2}, this error is the same as the computed error in Equation \ref{eq:ErrorSwarm1} but For the second algorithm \textsl{AlgS2} it is different than the computed error in Equation \ref{eq:errorSwarm2}. The additional computation is necessary to compare the two segmentations fairly. The system chooses the minimum segmentation as the final segmentation of this stroke.

After the best fit segmentation is chosen the parameter of each primitive is computed. For example, if the stroke is represented as an ellipse, the center and axis of the ellipse is computed and confirmed. For lines the slope, equation, start point and end point are computed.  These values and parameter will help determine the spatial relations between different segments in the feature extraction step. 
% The segmentation which crosseponds to the minimum $E_{tot}$ is the best segmentation.  
%\begin{equation}
%\label{eq:bestFit1}
%E_{area}=A_{st}-A_{sg}
% \end{equation}
% \begin{equation}
% where $A_{st}$ is the area of the input stroke, $A_{sg}$ is area of computed segmentation, $E_{alg}$ is the Segmentation Error.

@@

After the segmentation, each stroke is represented as a list of segments $L_s$. A segment can be an ellipse, a line or a circular arc. The list of segments $L_s$ represents the segmentation of the stroke. When a user draw another stroke the generated segmentation of the stroke is appended to the list of segments.  When the user finish drawing the symbol the segment list $L_s$ will contain all segments of the drawn symbol. The relation and properties between different segments are extracted from the segment list $L_s$.  This step is considered as feature extraction which is described in details in the next sections.

@@

\subsection{Structural and geometrical Features(FS1)}
 Features define the structure of the geometrical symbol. Section \ref{sec:steprec} describe in details how these 13 feature are computed.  
%\begin{description}
% \item [Structural and geometrical Features (FS1)] 
  \begin{itemize}
	 \item \emph{Segments:} Number of segments in the symbol.
	 \item \emph{Strokes:} Number of strokes or partial strokes that created the symbols.  
		\item  \emph{Primitives:} Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
		\item \emph{Curves:} Number of curves or ellipses in the symbol. 
		\item \emph{Lines:} Number of lines in the symbol. 
		\item \emph{Perpendicular} lines Number of perpendicular lines. 
		\item \emph{Parallel} lines Number of parallel lines. 
		\item \emph{Intersections} Number of intersection between lines and curves. 
		\item \emph{T intersections} Number of T intersections. 
		\item \emph{L intersections} Number of L intersections. 
		\item \emph{X intersections} Number of X intersections.
		\item \emph{Min Radius} Minimum Radius of all curves in the symbol.
		\item \emph{Max Radius} Maximum Radius of all curves in the symbol.
		%\item[No. of holes] Number of holes in the symbols. 
\end{itemize}
%\begin{description}
%	 \item[Segments] Number of segments in the symbol.
%	 \item [Strokes] Number of strokes or partial strokes that created the symbols.  
%		\item[Primitives] Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
%		\item [Curves] Number of curves or ellipses in the symbol. 
%		\item [Lines] Number of lines in the symbol. 
%		\item [Perpendicular lines] Number of perpendicular lines. 
%		\item [Parallel lines] Number of parallel lines. 
%		\item [Intersections] Number of intersection between lines and curves. 
%		\item [T intersections] Number of T intersections. 
%		\item [L intersections] Number of L intersections. 
%		\item [X intersections] Number of X intersections.
%		
%		
%		%\item[No. of holes] Number of holes in the symbols. 
%
%\end{description}
\subsection{Rubine Feature Set (FS2)} Features introduced by Rubine\cite{gestureexample12} for single stroke gestures. These 12 features represent the global shape of the symbol drawn. To compute the features we append all segments points into single path $ink_{path}$. The features is computed on the generated path $ink_{path}$.%then compute the features based on this path.  
\begin{enumerate}
	\item Cosine of starting angle.
	\item Sine of starting angle.
	\item Length of diagonal of bounding box. It gives an idea of the size of the bounding box).
	\item Angle of diagonal. It gives an idea of the shape of the bounding box (long, tall, square).
	\item Distance from start to end.  
	\item Total stroke length
	\item Change in Rotation( Arctan ). It gives the directional angle.
	\item Absolute rotation 
	\item Rotation squared 
	\item The maximum speed reached (squared) 
	\item Total time of stroke 
\end{enumerate}
\subsection{Statistical Features (FS3)} 
\begin{enumerate}
	\item \textsl{Zernike moments:} Zernike moments of order $n$\footnote{ We tried different values of n. Results show the analysis of each order.} \cite{HeloiseBeautification}.  For $n=10$ there are 32 moment features. 
\end{enumerate}
\subsection{Global shape properties set (FS4)}
 Features composed of different geometrical and statistical properties. These 13 features represent the global shape of the symbol drawn.

@@

\section{Classification}%%Feature Set
\label{sec:Classification}
The system uses support vector machine (SVM) classifier with Linear kernel (Linear kernel) \cite{libsvm}. SVM is a binary classifier that maximizes the margin between the decision boundary and the support vectors. To enable SVM to learn multi classes we used one versus one strategy for combining set of binary classifier. This strategy trains $ \frac{n(n-1)}{{2}}$ binary classifier each distinguish between only two shapes. The final result is based on a number of votes each shape has from all classifiers. We used the library implementation of LibSVM \cite{libsvm} and its implementation of the one-vs-one strategy. Section \ref{Sec:SVMdetail} describe in details the classifier, how classes (classification) are determined and the reasons behind using this specific classification method.

@@

\subsection{Support Vector Machines}
\label{Sec:SVMdetail}
The foundation of Support Vector Machines (SVM) were developed by Vapnik \cite{svmintroduce} in 1995. It is gaining increase popularity due to the high performance and efficient implementation. The goal of SVM algorithms is to generate a separating plane between two different classes which are represented using a set of n-dimensional vectors. The generated plane is used as a good classifier to the unseen examples. However, there can be more than one plane that can separates the data therefore SVM attempts to find the plane that maximize the margin (distance between classifier and nearest sample) between the two classes \ref{fig:svm1}. To calculate the margin, two different hyper planes are constructed, one plane on each side of the data, these two planes are then moved to generate the maximum margin possible.  
\begin{figure}
	\centering
		\includegraphics{afterDefense/svm1.jpg.eps}
	\caption{Different Classifier Planes}
	\label{fig:svm1}
\end{figure}

@@

 The orientation of the separation plane is determined using the vector $\omega$ and the offset of the plane from the origin is determined using the scalar $b$.   Let's make an assumption that the data points are linearly separable, i.e. there are infinite possibilities to find a plane that separates and correctly classify the training data. This is idea is illustrated in Figure\ref{fig:svm1} where two different separating planes can correctly classify the data. It is intuitively noted from the figure that the solid line is more likely to generalize more data. This means less classification error will be introduced. The geometrical characterization of this plane means the plane furthest from both classes.     % Which one is preferable ? Intuitively one prefers the solid pane since small perturbations of any point would not introduce misclassifications errors. Without any information, the solid plane is more likely to generalize better on future data. Geometrically we can characterize the solid plane as being furtherer from both classes.  Plane supports a class if all points in that class are on one side of that plane.

@@

\subsubsection{Linearly Inseparable Case}
In all previous sections we assumed that the problem of classifying two set is linearly separable. The assumption is not true in most cases, thus, trying to construct a plane to separate the two dataset will fail. Figure \ref{fig:svm5} illustrates that we cannot construct a plane that can separate the two sets correctly.  Constrains for constructing the planes must be relaxed since the QP task is not feasible for the linearly inseparable case.  Consider the linear inseparable problem in Figure\ref{fig:svm5}, in the ideal case there will be no point misclassified and no point falls on the margin. But we must relax constrains that insure that each point in on the appropriate side of it is supporting plane. Every point that falls on the wrong side of the supporting plane is considered an error. The problem is now to maximize the margin and minimize the error.

@@

\subsubsection {Non linear function Using Kernels}
Figure \ref{fig:svm6} shows a classification problem that cannot be separated using linear space. A higher order plane can work better in similar cases. Converting a linear classification algorithm into non linear classification algorithm can be achieved by adding additional attribute to the data that are nonlinear functions of the original data. The expanded dataset features space can be separated using an existing linear classification algorithm.  The linear classification that is applied on the feature space produces a non linear classification on the original input space. In two dimensional vector space with attributes $r$ and $s$, to construct a quadratic  discriminate  simply map the original two dimensional input space $\left[r,s\right]$ to  five dimensional features space $\left[r,s,rs,r^2,s^2\right]$ and construct a linear discriminate in that space. Specifically, define : $\vartheta(x):R^2  \rightarrow R^5$ then 
\begin{equation}
\begin{array}{l}
	x=[r,s] \\
	\omega \cdot x=\omega_1 r+ \omega_2 s \\
	\downarrow \\
	\vartheta(x)=\left[r,s,rs,r^2,s^2\right] \\

@@

 There is mainly two popular methods to change the binary SVM classification problem into Multi class classification problem. The first method is One Versus All (OVA) Strategy where M binary classifiers is constructed one classifier for each class.  The $ith$ classifier is trained using positive example form $ith$ class and negative example from all other classes. The output function $p_i$ represents the distance and classification of the example from the $ith$ classifier. For a new example $x$, OVA SVM strategy assigns it to the class with the largest value of $p_i$.

@@

SVM have been applied to the a largest MNIST a Handwritten digit recognition dataset of 60000 training set and 10000 test set\cite{IjdarSherifPaper}. Sherif and Ezzat has performed a comparison between SVM and several other techniques on the MNIST dataset. They also compared the same classification methods on Arabic handwritten digits (MADBASE) \cite{ADBase9,IjdarSherifPaper}. Their result shows that SVM classifier outperforms all other techniques on both MNIST and MADBase datasets and achieved recognition rate of over 98\%. Similar results was reported by \cite{empiricalcomp11,SVMInvariantComDecoste02} on the MNIST dataset. There are many other successful applications.  SVM have proven to be robust to noise and perform well on many tasks \cite{empiricalcomp11,Scholkopf97comparingsupport}.

This superior performance because they may eliminates many of problems experienced with other inference methodologies with other inference methodologies like neural network and decision trees. The first problem that is eliminated was the problems with local minima. SVM can construct highly nonlinear classification and regression function without working about getting stuck at local minima. Another problem SVM eliminated is the numbers of parameter that are picked to generate the model. For example, SVM classifier with RBF kernel will only need 2 parameters to set the model the $C$ cost parameter and $\sigma$ parameter of the RBF kernel. The last reason is the final result are sable, reproducible and largely independent of the specific algorithm used to optimize the SVM model. If two users apply the same SVM model with the same parameters to the same data, they will get the same solution module numeric issues. Compare this with neural networks where the results are dependent on the particular algorithms and starting point used.

@@

\subsubsection {SVM Implementations}
To solve SVM approach requires solving a QP or LP problem.  Mathematical programming has extensively studied the LP and QP type problems. These techniques can be divided into three categories, methods in which evaluate and discard kernel components during learning. Decomposition methods in which evolving subsets of data are used and new optimization approaches that specifically utilize the structure of the SVM problem.

@@

The third approach is to directly attack the SVM problem from and optimization perspective and create an algorithm that explicitly exploits the structure of ten problems. Keerthi et al. \cite{fastSVM6} proposed a very effective algorithm based on the dual geometry of finding the two closest points in the convex hulls.

@@

In this step the user draw one or more stroke to draw a symbol. Figure \ref{fig:clock} illustrate the sequence and direction of user draw. In this section, each stroke user draws will be labeled as it illustrated in Figure \ref{fig:clockstroke1} and Figure \ref{fig:clockstroke2}.  Table \ref{tab:StepsStroke1} shows the detailed output of the program at each step. The input of this step is a set of strokes, each stroke has a set of points and the output is a list of segments that will be used in the recognition process. More detailed example and algorithms are demonstrated in the Appendix \ref{ChapterstepExample} \\

@@

%	\includegraphics[scale=0.5]{afterDefense/clockstroke1.eps}
  &   &    \includegraphics[scale=0.5]{afterDefense/stroke1withpdp.jpg.eps}
\\ \hline 
Segmentation & Segmentation of Ellipse & $P_{dp}$   &  Ellipse fitting  & 
The circle is   a = 103.05715108084719 ,b = 103.06851874747797 , Center( 283.03365107691775,  181.05035287667172 )   with Error =  1.9471676673301757 , Percent =1.0695193086531243
eff  = 0.27463462099275116

@@

Stroke 2 & Extracting point and computing boundary data & -- &  see Section \ref{sec:Preprocessing} &
 	\includegraphics[scale=0.5]{afterDefense/clockstroke2.jpg.eps} \\ \hline
  & Possible Dominate point Extraction for details see Section\ref{sec:Preprocessing} &   	%\includegraphics[scale=0.5]{afterDefense/clockstroke2.jpg.eps}
    & 1. compute the $P_{pd}$  &  
  \includegraphics[scale=0.5]{afterDefense/stroke2WithPdp.jpg.eps}
    \\ \hline
